Calculating housing and factor shares from industry data

### 2023-01-18 PhD_pages/data_appendices/1_A_prepare_PIP_and_WID.qmd
restructuring and data work on paper 1

@@

#  A local version, in case internet isn't available 
df_pip<- read.csv("PhD_pages/data/pip_dataset (2).csv")

@@

Here we define a list of these aggregate entities and drop them from the data.
```{r}
drop_aggs<- c(
  "East Asia and Pacific",
  "Europe and Central Asia",
  "High income countries",
  "Latin America and the Caribbean",
  "Middle East and North Africa",
  "Sub-Saharan Africa",
  "World"
)

@@

@@

df_pip<- df_pip %>%
  rename("Gini" = "gini",
         "Top_10_share" = "decile10_share",
         "MLD" = "mld",
         "P90_P10_ratio" = "p90_p10_ratio",
         "Entity" = "country",
         "Year" = "year")

df_pip<- df_pip %>%
    select(
        Year,
        Entity,
        Gini,
        Top_10_share,
        MLD,
        P90_P10_ratio,
        welfare_type)

@@

@@

# A local version, in case internet isn't available 
df_wid<- read.csv("PhD_pages/data/wid.csv")

@@

Here we create a list of entities in the WID data that do not correspond to countries in existence today. These we drop from the data.
```{r}
drop_entities<- c(
"China - rural",
"China - urban",
"East Germany",
"Other Russia and Central Asia",
"Other East Asia",
"Other Western Europe",
"Other Latin America",
"Other MENA",
"Other South & South-East Asia",
"Other Sub-Saharan Africa",
"Africa",
"Asia",
"Europe",
"Oceania",
"Central Asia",
"East Africa",
"East Asia",
"Eastern Europe",
"Middle Africa",
"North Africa",
"North America",
"South-East Asia",
"South Africa region",
"West Africa",
"West Asia",
"Western Europe",
"European Union",
"World",
"Asia (excluding Middle East)",
"North America and Oceania",
"Sub-Saharan Africa",
"Latin America",
"Middle East",
"MENA",
"Russia and Central Asia",
"South & South-East Asia"
)

@@

@@

```{r}
df_wid<- df_wid %>%
    rename("Gini" = "Gini.coefficient",
         "Top_10_share" = "P90.P100...share.of.the.top.10.") %>%
    select(
        Year,
        Entity,
        Gini,
        Top_10_share,
        # MLD,
        P90_P10_ratio)

@@

@@

>>>>

```{python}
# For some reason I can't get the owid package working when rendering... only when running in the interactive window in vs code
# from owid import catalog
import pandas as pd
```
:::

@@

#  A local version, in case internet isn't available 
df_pip<- read.csv("PhD_pages/data_appendices/data/original/pip_dataset.csv")

@@

Here we define a list of these aggregate entities and drop them from the data.
```{r}
drop_aggs<- c(
  "East Asia and Pacific",
  "South Asia",
  "Europe and Central Asia",
  "High income countries",
  "Latin America and the Caribbean",
  "Middle East and North Africa",
  "Sub-Saharan Africa",
  "World"
)

@@

::: {.callout-note collapse="true" appearance="minimal"}
### Drop extremely low P10 values

There are some very low values of P10. These lead to exaggerated values of the P90/P10 ratio.

kable(
    df_pip %>% filter(decile1_thr<0.02)

We set these values of P10 to NA, and thereby precluding a value for P90/P10 ratio when it is calculated below.

df_pip<- df_pip %>%
  mutate(decile1_thr = if_else(decile1_thr<0.02,
                              as.numeric(NA),
                              decile1_thr))

kable(
    df_pip %>% filter(is.na(decile1_thr))

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Calculate additional variables

df_pip<- df_pip %>%
  mutate(Bottom_50_share = decile1_share + decile2_share + decile3_share + decile4_share + decile5_share,
          P90_P10_ratio = decile9_thr/decile1_thr)

@@

df_pip<- df_pip %>%
  rename("Gini" = "gini",
         "Top_10_share" = "decile10_share",
         "MLD" = "mld",
         "Entity" = "country",
         "Year" = "year")

df_pip<- df_pip %>%
    select(
        Year,
        Entity,
        Gini,
        Top_10_share,
        Bottom_50_share,
        MLD,
        P90_P10_ratio,
        welfare_type)

@@

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Save the data
I save the prepared data locally:
```{r}

write.csv(df_pip, 
      "PhD_pages/data_appendices/data/clean/pip.csv", row.names=FALSE)

@@

# A local version, in case internet isn't available 
df_wid<- read.csv("PhD_pages/data_appendices/data/original/wid.csv")

@@

Here we create a list of entities in the WID data that do not correspond to countries in existence today. These we drop from the data.
```{r}
drop_entities<- c(
"China - rural",
"China - urban",
"East Germany",
"Other Russia and Central Asia",
"Other East Asia",
"Other Western Europe",
"Other Latin America",
"Other MENA",
"Other South & South-East Asia",
"Other Sub-Saharan Africa",
"Africa",
"Asia",
"Europe",
"Oceania",
"Central Asia",
"East Africa",
"East Asia",
"South Asia",
"Eastern Europe",
"Middle Africa",
"North Africa",
"North America",
"South-East Asia",
"South Africa region",
"West Africa",
"West Asia",
"Western Europe",
"European Union",
"World",
"Asia (excluding Middle East)",
"North America and Oceania",
"Sub-Saharan Africa",
"Latin America",
"Middle East",
"MENA",
"Russia and Central Asia",
"South & South-East Asia"
)

@@

::: {.callout-note collapse="true" appearance="minimal"}
### Drop extremely low P10 values

There are some very low values of P10. These lead to exaggerated values of the P90/P10 ratio. Here are all values less than $1 a day (using pretax national income)

kable(
    df_wid %>% filter(P10...income.threshold/365<1)

We set these values of P10 to NA, and thereby precluding a value for P90/P10 ratio when it is calculated below.

df_wid<- df_wid %>%
  mutate(P10...income.threshold = if_else(P10...income.threshold/365<1,
                              as.numeric(NA),
                              P10...income.threshold))

                              kable(
    df_wid %>% filter(is.na(P10...income.threshold))

:::

@@

```{r}
df_wid<- df_wid %>%
    rename("Gini" = "Gini.coefficient",
         "Top_10_share" = "P90.P100...share.of.the.top.10.",
         "Bottom_50_share"= "P0.P50...share.of.the.bottom.50.",
         "P90" = "P90...income.threshold.of.the.top.10.",
         "P10" = "P10...income.threshold") %>%
    select(
        Year,
        Entity,
        Gini,
        Top_10_share,
        # MLD,
        P90_P10_ratio,
        Bottom_50_share,
        P90,
        P10)

@@

::: {.callout-note collapse="true" appearance="minimal"}
### Save the data
I save the prepared data locally:
```{r}

write.csv(df_wid, 
      "PhD_pages/data_appendices/data/clean/wid.csv", row.names=FALSE)

```
:::
:::

## Additional variables
::: {.callout-note collapse="true" appearance="minimal"}
### Region classfications
For regional classifications I take WID definitions, as specified on their website [here](https://wid.world/codes-dictionary/#country-code).

I read in a csv prepared from the table in that webpage, which has been mapped to the Our World in Data standarized country names used throughout this project.
```{r}

df_regions<- read.csv("PhD_pages/data_appendices/data/original/WID_regions_country_standardized.csv")

I create a clean region mapping table using two regional definitions. One is the 'standard' definitions used by WID. In an alternative set I split Western and Eastern Europe, ccombining the latter with Russia & Central Asia to make a joint Eastern Europe and Central Asia region. 
```{r}

df_regions<- df_regions %>%
  select(Our.World.In.Data.Name, `region..standard.`, sub.region) %>%
  rename(Entity = Our.World.In.Data.Name,
        WID_region = `region..standard.`) %>%
  mutate(region_alt = if_else(sub.region == 'Western Europe',
  'Western Europe',
  WID_region)) %>%
  mutate(region_alt = if_else((sub.region == 'Eastern Europe' | WID_region == "Russia & Central Asia"),
  'Eastern Europe & Central Asia',
  WID_region)) %>%
  select(Entity, WID_region, region_alt)

kable(
  df_regions
)
```

Check that no countries are not matched for either dataset.
```{r}
# Merge in the regions dataset
df_pip<- left_join(df_pip, df_regions)
df_wid<- left_join(df_wid, df_regions)

# PIP check
kable(
  df_pip %>% 
    select(Entity, WID_region, region_alt) %>%
    unique() %>%
    filter(is.na(WID_region))
)

# WID check
kable(
  df_wid %>% 
    select(Entity, WID_region, region_alt) %>%
    unique() %>%
    filter(is.na(WID_region))
)

Save the region mapping
```{r}
write.csv(df_regions, 
      "PhD_pages/data_appendices/data/clean/WID_region_mapping.csv", row.names=FALSE)

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Region classfications

```{python}
# For some reason I can't get the owid package working when rendering... only when running in the interactive window in vs code
# results = catalog.find('population')
# owid_pop = results[results['path'] == 'garden/owid/latest/key_indicators/population']

# df_pop = owid_pop.iloc[0].load().reset_index().rename(columns={'year':'Year', 'country': 'Entity'})

# df_pop.head()

Save the population data
```{python}
# df_pop.to_csv('PhD_pages/data_appendices/data/clean/population.csv')

@@

### Outliers
::: {.callout-note collapse="true" appearance="minimal"}
### P90/P10 ratio

For the P90/P10 ratio, a few observations have extreme values. In the PIP data
```{r}
kable(

  df_pip %>% arrange(-P90_P10_ratio) %>% head()
)

And the WID data:

```{r}
kable(

  df_wid %>% arrange(-P90_P10_ratio) %>% head(100)
)

:::

### 2023-01-18 PhD_pages/data_appendices/1_B_within_country_trends.qmd
restructuring and data work on paper 1

---
title: "Appendix 1.B: Analysing global trends in within-country inequality"
format: html
warning: false
filters:
  - shinylive
---

@@

>>>>

---
title: "Appendix 1.B: Analysing global trends in within-country inequality"
format: html
warning: false
filters:
  - shinylive
---
```{python}

import pandas as pd
import numpy as np
from plotnine import *
import plotly.graph_objects as go
```

Load in the data
```{python}

df_wid = pd.read_csv("data/clean/wid.csv")
df_pip = pd.read_csv("data/clean/pip.csv")
df_regions = pd.read_csv("data/clean/WID_region_mapping.csv")

```{python}

# df_wid = pd.read_csv("PhD_pages/data_appendices/data/clean/wid.csv")
# df_pip = pd.read_csv("PhD_pages/data_appendices/data/clean/pip.csv")
# df_regions = pd.read_csv("PhD_pages/data_appendices/data/clean/WID_region_mapping.csv")

@@

### Mapping data to a reference year
This function grabs the observation, by a grouping variable (e.g. by country), for which a reference variable (e.g. year) is closest to a reference value (e.g. 2002).
You can specify a maximum distance from the reference value, beyond which no match will be returned.
Because there may be tie-breaks – matches equally distant above or below the reference value – there is an argmuent to specify how these tie-breaks should resolved.
```{python}

#  Function get matching for ref years
def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):

    df = df.loc[:, [reference_col, group_by_col, value_col]]

    # Drop NAs
    df = df.dropna()

    # Calculate absolute distance from reference value
    df['ref_diff'] = abs(df[reference_col] - reference_val)

    # Drop any rows with a distance beyond threshold
    if not pd.isna(max_dist_from_ref):
      df = df.loc[df['ref_diff'] <= max_dist_from_ref]

    # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
    df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)

    # Settle tie-breaks
    if tie_break == 'below':
      df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)

    elif tie_break == 'above':
      df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)

    df = df.drop('ref_diff', axis=1)

    return df

I test this function here:
```{python}

test = closest_to_reference(
  df = df_wid, 
  reference_val = 2002,
  max_dist_from_ref = 5,
  reference_col = 'Year',
  group_by_col = 'Entity',
  value_col = 'Gini',
  tie_break = 'below'
  )

test.head()

### Obtaining a pair of observations

This function runs the `closest_to_reference` function twice, over a pair of reference values.
Tie-breaks are settled so as to maximise the gap between the two reference values.
You can also specify a minimum distance between the two observations (e.g. pairs of matches that fall less than X years apart will be dropped).

```{python}
# Merge matches for different reference points
def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

  # Make sure the pair of reference values are in ascending order
  reference_vals.sort()

  # Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref

  # Find matches for lower reference value
  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs[0], reference_col, group_by_col, value_col, 'below')

  # Find matches for higher reference value
  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs[1], reference_col, group_by_col, value_col, 'above')

  # Merge the two sets of matches
  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))

  # Drop obs that do not have data for both ref values
  merged_df = merged_df.dropna()

  # Drop obs where the matched data does not meet the min distance requirement
  if not pd.isna(min_dist_between):

    # Store the names of the reference column returned from the two matches
    ref_var_high = f'{reference_col}{reference_vals[1]}'
    ref_var_low = f'{reference_col}{reference_vals[0]}'

    # Keep only rows >= to the min distance
    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]

  return merged_df

I test this function here:
```{python}
test = merge_two_ref_matches(
  df = df_wid, 
  reference_vals = [2000, 2010], 
  max_dist_from_refs = [5, 4],
  min_dist_between = 9,
  reference_col = 'Year',
  group_by_col = 'Entity',
  value_col = 'Gini'
  )

test.head()

#### Allowing for the different welfare concepts in the PIP data

As noted in .... (add backlink) ... the PIP data contains both income and consumption observations.
This function produces matched pairs of observations from that data in which only one pair is selected for each value of the grouping variable (i.e. each country).
Priority is given to pairs for which both observations relate to income. Second priority is given to pairs where both observations relate to consumption.
Only if neither pair is available then it will return a mixed pair of observations if that is available.
```{python}
# For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):

  # Specify the name of the column in which the income/consumption welfare definition is stored
  welfare_colname = 'welfare_type'

  # Creat dataframes for thee scenarios:
  # Scenario 1: only allow income data
  df_inc_filter = df.loc[df[welfare_colname] == "income", :]
  df_inc_filter.name = "Income"

  # Scenario 2: only allow consumption data
  df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
  df_cons_filter.name = "Consumption"
  # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
  df_mixed = df.copy()

  df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')

  df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]

  df_mixed.name = "Mixed"
  #  Store the scneario dataframes in a list
  df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]

  # Run the matching function on each scenario
  scenario_matches = [merge_two_ref_matches(
    df_scenario, 
    reference_vals, 
    max_dist_from_refs, 
    min_dist_between, 
    reference_col, 
    group_by_col, 
    value_col) for df_scenario in df_scenarios]

  # Combine the first two scenarios.
  df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])

  # Tidy up indexes
  df_combined_matches = df_combined_matches.reset_index()

  df_combined_matches = df_combined_matches.drop('level_1', axis=1)

  df_combined_matches = df_combined_matches\
    .rename(columns={"level_0": "pip_welfare"})

  # Add in third scenario.
  df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])

  # add scenario name to te pip_welfare column
  df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)

  # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
    # First count the matches
  df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
    # Then drop any matches from the lowest priority where there are multiple matches
  df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
    #  Repeat at the next level of priority
  df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
  df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]

  # Drop the match count column
  df_combined_matches = df_combined_matches.drop('match_count', axis=1)

  return df_combined_matches

I test this function here:
```{python}

test = pip_welfare_routine(

  df = df_pip,
  reference_vals = [1986, 2016], 
  max_dist_from_refs = [5, 5], 
  min_dist_between = 30, 
  reference_col = 'Year',
  group_by_col = 'Entity',
  value_col = 'Gini'

  )

test.head()
```

### Merging reference year aligned data from WID and PIP datasets

```{python}

def prep_wid_pip_data(reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
  pip_matches = pip_welfare_routine(
    df = df_pip,
    reference_vals = reference_vals,
    max_dist_from_refs = max_dist_from_refs,
    min_dist_between = min_dist_between,
    reference_col = reference_col,
    group_by_col = group_by_col,
    value_col = value_col
  )

  wid_matches = merge_two_ref_matches(
    df = df_wid,
    reference_vals = reference_vals,
    max_dist_from_refs = max_dist_from_refs,
    min_dist_between = min_dist_between,
    reference_col = reference_col,
    group_by_col = group_by_col,
    value_col = value_col
  )

  ref_pairs = pd.concat([pip_matches, wid_matches,], keys=['pip', 'wid'])

  # Tidy up indexes
  ref_pairs = ref_pairs.reset_index()

  ref_pairs = ref_pairs.drop('level_1', axis=1)

  ref_pairs = ref_pairs\
    .rename(columns={"level_0": "source"})

  return ref_pairs
```

## Prepare plot and tables

```{python}
#| echo: false

def prep_plot_and_tables(
  reference_vals, 
  max_dist_from_refs,
  min_dist_between,
  reference_col,
  group_by_col,
  value_col,
  tolerance,
  outlier_cut_off_upper,
  source_count_requirement):

  # Other spec:
  region_col = 'region_alt'

  # ---- Prep data ---------

  plot_data = prep_wid_pip_data(
    reference_vals = reference_vals, 
    max_dist_from_refs = max_dist_from_refs,
    min_dist_between = min_dist_between,
    reference_col = reference_col,
    group_by_col = group_by_col,
    value_col = value_col,
  )

  # Store the names of the columns to be used onthe X and Y axis
  x_axis = f'{value_col}{reference_vals[0]}'
  y_axis = f'{value_col}{reference_vals[1]}'

  # Apply outlier cut off, if specified
  if not pd.isna(outlier_cut_off_upper):
    plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
    plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]

  # Add a count by country – showing whether data is available from both sources or not
  plot_data['source_count'] = plot_data.groupby(group_by_col)['source'].transform('count')

  # Apply source requirement (whether to include only observations with data from both sources)
  plot_data = plot_data.loc[plot_data['source_count'] >= source_count_requirement]

  # Drop source_count column
  plot_data = plot_data.drop('source_count', axis=1)

  # Add in region classification
  plot_data = pd.merge(plot_data, df_regions, how = 'left')

  # Make nicely formatted table
  plot_data_formatted = plot_data.copy()

  plot_data_formatted[x_axis] = plot_data_formatted[x_axis].map('{:.2f}'.format)
  plot_data_formatted[y_axis] = plot_data_formatted[y_axis].map('{:.2f}'.format)

  plot_data_table = go.Figure(data=[go.Table(
    header=dict(values=list(plot_data_formatted.columns)),
    cells=dict(values=plot_data_formatted.transpose().values.tolist() ))
    ])

  # ----- Prep plot ------

  # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
  x_min = plot_data[x_axis].min()
  x_min_floor = np.floor(x_min * 10) / 10

  x_max = plot_data[x_axis].max()
  x_max_ceiling = np.ceil(x_max * 10) / 10

  # Set the coordinates of the shaded ribbon
  shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
    'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})

  plot = (ggplot(plot_data
  , aes(x_axis, y_axis))
  + geom_point(aes(color=region_col))
  + facet_wrap('~ source')
  + geom_polygon(aes(x='x', y='y'), data=shaded_coord, fill="#FF000066"))

  # ---- Prep summary table ----

  # Count by rise/stable/fall bins

  # Calculate the change between the two ref periods
  plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])

  # Define the thresholds using the tolerance specified
  bin_thresholds = [-float("inf"), -tolerance, tolerance, float("inf")]
  bin_labels = ['fall', 'stable', 'rise']

  bins = pd.cut(plot_data['change'], bins=bin_thresholds, labels=bin_labels)

  # By region:

  # Count observations in bins defined by the thresholds
  threshold_counts = plot_data\
    .groupby(['source', region_col, bins])\
    .size()\
    .unstack()\
    .unstack(level=region_col)\
    .T\
    .reset_index()\
    .rename(columns={"change": "summary"})

  # Calculate average change and n
  mean_and_count = pd.DataFrame(plot_data.groupby(['source', region_col])['change']\
    .agg(['mean', 'count']))\
    .unstack()\
    .T\
    .reset_index()\
    .rename(columns={"level_0": "summary"})

  # rename the labels coming out of the aggregation
  mean_and_count['summary'] = mean_and_count['summary'].replace(['mean', 'count'], ['avg change', 'n'])

  # Concat together and present in a table
  df_summary_region = pd.concat([threshold_counts, mean_and_count])

  df_summary_region['summary'] = pd.Categorical(df_summary_region['summary'], categories=['fall', 'stable', 'rise','n', 'avg change'], ordered=True)

  df_summary_region = df_summary_region.sort_values([region_col,'summary'])

  # Set the index
  df_summary_region = df_summary_region.set_index([region_col,'summary'])

  # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
  repeat_length = 5
  for i in range(0, len(df_summary_region.index.levels[0].unique())):
    df_summary_region.iloc[i*repeat_length,:] = df_summary_region.iloc[i*repeat_length,:].map('{:.0f}'.format)
    df_summary_region.iloc[1 + i*repeat_length,:] = df_summary_region.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary_region.iloc[2 + i*repeat_length,:] = df_summary_region.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary_region.iloc[3 + i*repeat_length,:] = df_summary_region.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary_region.iloc[4 + i*repeat_length,:] = df_summary_region.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)

  # World summary

  # Count observations in bins defined by the thresholds
  threshold_counts = plot_data\
    .groupby(['source', bins])\
    .size()\
    .unstack(level='source')\
    .reset_index()\
    .rename(columns={"change": "summary"})

  # Calculate average change and n
  mean_and_count = pd.DataFrame(plot_data.groupby(['source'])['change']\
    .agg(['mean', 'count']))\
    .T\
    .reset_index()\
    .rename(columns={"index": "summary"})

  # rename the labels coming out of the aggregation
  mean_and_count['summary'] = mean_and_count['summary'].replace(['mean', 'count'], ['avg change', 'n'])

  # Concat together and present in a table
  df_summary_world = pd.concat([threshold_counts, mean_and_count])

  df_summary_world['summary'] = pd.Categorical(df_summary_world['summary'], categories=['fall', 'stable', 'rise','n', 'avg change'], ordered=True)

  df_summary_world = df_summary_world.sort_values(['summary']).set_index('summary')

  # format to show appropriate d.p. I do it row by row for want of figuring out a better way!
  df_summary_world.iloc[0,:] = df_summary_world.iloc[0,:].map('{:.0f}'.format)
  df_summary_world.iloc[1,:] = df_summary_world.iloc[1,:].map('{:.0f}'.format)
  df_summary_world.iloc[2,:] = df_summary_world.iloc[2,:].map('{:.0f}'.format)
  df_summary_world.iloc[3,:] = df_summary_world.iloc[3,:].map('{:.0f}'.format)
  df_summary_world.iloc[4,:] = df_summary_world.iloc[4,:].map('{:.2f}'.format)

  # ----- Return outputs ------
  plot_and_tables = {
    "table_data": plot_data,
    "table": plot_data_table,
    "plot": plot,
    "summary_world": df_summary_world,
    "summary_region": df_summary_region
  }

  return plot_and_tables

```{python}

# Specifications
reference_vals = [1990, 2018]
max_dist_from_refs = [4, 4]
min_dist_between = 20
value_col = 'Bottom_50_share'
# 'Gini'
# 'Top_10_share'
# 'Bottom_50_share'
# 'MLD'
# 'P90_P10_ratio' – not that helpful due to accuracy of low incomes

# Data from both sources is needed?
  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
  # If = 2, then only countries with data for both sources will be included
source_count_requirement = 2

# Set the value which constitutes a significant change in the value being measured
  # E.g. for Gini tolerance = 0.02 (2 points)

if value_col == 'Gini':
  tolerance = 0.02
  outlier_cut_off_upper = None

if value_col == 'Top_10_share':
  tolerance = 1
  outlier_cut_off_upper = None

if value_col == 'Bottom_50_share':
  tolerance = 1
  outlier_cut_off_upper = None

if value_col == 'P90_P10_ratio':
  tolerance = 0.5
  outlier_cut_off_upper = 40

# These parameters are unlikely to change:
reference_col = 'Year'
group_by_col = 'Entity'

plot_and_tables = prep_plot_and_tables(
  reference_vals = reference_vals, 
  max_dist_from_refs = max_dist_from_refs,
  min_dist_between = min_dist_between,
  reference_col = reference_col,
  group_by_col = group_by_col,
  value_col = value_col,
  tolerance = tolerance,
  outlier_cut_off_upper = outlier_cut_off_upper,
  source_count_requirement = source_count_requirement
  )

```{python}
plot_and_tables['plot']
```

```{python}
plot_and_tables['summary']
```

:::{.column-page}
```{python}

plot_and_tables['table'].show()

```{python}

plot_data_results = plot_and_tables['table_data']

```
:::

### 2023-01-18 PhD_pages/data_appendices/3_C_calc_factor_shares.qmd
restructuring and data work on paper 1

---
title: "Calculating housing and factor shares from industry data"
editor: visual
---

>>>>

---
title: "Calculating housing and factor shares from industry data"
editor: visual
execute:
  freeze: true  # never re-render during project render
---

# To-do:

- I moved the data folder, but haven't yet relected this in the code below.
